---
title: "ICE ERO-LESA enforcement actions data per AOR, FY12-22"
author: "UWCHR"
date: "2024-06-06"
params:
  aor: SEA
output:
    html_document:
        html_preview: true
        toc: true
        toc_depth: 3
        toc_float: true
        code_folding: hide
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}

library(pacman)

p_load(here, tidyverse, zoo, lubridate, ggplot2, plotly, gghighlight, sf, tigris, ggrepel, viridis)

options(scipen = 1000000)

specific_aor <- params$aor

```

# ICE ERO enforcement data: `r specific_aor` AOR

```{r load_data, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}

# ENCOUNTERS DATA

enc <- read_delim(here('analyze', 'input', 'ice_encounters_fy12-23ytd.csv.gz'), delim='|',
                  col_types = cols(aor = col_factor(),
                                   event_date = col_character(),
                                   landmark = col_character(),
                                   operation = col_factor(),
                                   processing_disposition = col_factor(),
                                   citizenship_country = col_factor(),
                                   gender = col_factor(),
                                   hashid = col_character(),
                                   id = col_number()))

# glimpse(enc)

redacted <- c('encounter_threat_level', 'alien_file_number')

enc <- enc %>% 
  dplyr::select(-redacted)

enc <- enc %>% 
  mutate(aor = factor(aor, levels = sort(levels(enc$aor))),
         event_date = as_date(event_date, format="%m/%d/%Y"),
         year = year(event_date),
         month = month(event_date, label=TRUE, abbr=TRUE),
         year_mth = zoo::as.yearmon(event_date),
         fy_quarter = as.factor(quarter(event_date, fiscal_start=10, type="year.quarter")),
         fy = as.factor(substr(fy_quarter, 1,4)),
         gender = toupper(gender),
         operation = toupper(operation),
         processing_disposition = toupper(processing_disposition),
         citizenship_country = toupper(citizenship_country)) %>% 
  filter(event_date >= "2011-10-01",
         event_date <= "2022-09-30")

# ARRESTS DATA

arr <- read_delim(here('analyze', 'input', 'ice_arrests_fy12-23ytd.csv.gz'), delim='|',
                  col_types = cols(aor = col_factor(),
                                  arrest_date = col_date(format="%m/%d/%Y"),
                                  departed_date = col_date(format="%m/%d/%Y"),
                                  apprehension_landmark = col_factor(),
                                  arrest_method = col_factor(),
                                  operation = col_factor(),
                                  processing_disposition = col_factor(),
                                  citizenship_country = col_factor(),
                                  gender = col_factor(),
                                  case_closed_date = col_date(format="%m/%d/%Y"),
                                  id = col_integer(),
                                  hashid = col_character()
                                  )) 

redacted <- c('removal_threat_level', 'apprehension_threat_level', 'alien_file_number')

arr <- arr %>% 
  dplyr::select(-all_of(redacted))

arr <- arr %>% 
  mutate(aor = factor(aor, levels = sort(levels(arr$aor))),
         arrest_date = as_date(arrest_date, format="%m/%d/%Y"),
         year = year(arrest_date),
         month = month(arrest_date, label=TRUE, abbr=TRUE),
         year_mth = zoo::as.yearmon(arrest_date),
         fy_quarter = as.factor(quarter(arrest_date, fiscal_start=10, type="year.quarter")),
         fy = as.factor(substr(fy_quarter, 1,4)),
         citizenship_country = as.factor(toupper(citizenship_country))) %>% 
  filter(arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30")

# REMOVALS DATA

pd_dict <- read_delim(here('share', 'hand', 'processing_disp.csv'), delim='|')

rem <- read_delim(here('analyze', 'input', 'ice_removals_fy12-23ytd.csv.gz'), delim='|',
                  col_types = cols(aor = col_factor(),
                                   arrest_date = col_date(format="%m/%d/%Y"),
                                  departed_date = col_date(format="%m/%d/%Y"),
                                  case_close_date = col_date(format="%m/%d/%Y"),
                                  removal_date = col_date(format="%m/%d/%Y"),
                                  apprehension_method_code = col_character(),
                                  processing_disposition_code = col_factor(),
                                  citizenship_country = col_factor(),
                                  gender = col_factor(),
                                  final_charge_section = col_factor(),
                                  id = col_integer(),
                                  hashid = col_character()
                                  )) 

redacted <- c('removal_threat_level', 'alien_file_number')

rem <- rem %>% 
  dplyr::select(-redacted, -case_closed_date)

rem <- rem %>% 
  mutate(aor = factor(aor, levels = sort(levels(rem$aor))),
         year = year(departed_date),
         month = month(departed_date, label=TRUE, abbr=TRUE),
         year_mth = zoo::as.yearmon(departed_date),
         processing_disp = toupper(coalesce(processing_disposition_code, processing_disposition)),
         fy_quarter = as.factor(quarter(departed_date, fiscal_start=10, type="year.quarter")),
         fy = as.factor(substr(fy_quarter, 1,4))) %>% 
  filter(departed_date >= "2011-10-01",
         departed_date <= "2022-09-30")

rem <- left_join(rem, pd_dict, by=c('processing_disp' = 'processing_disposition_raw'))

# SUPPLEMENTAL DATA

demog <- read_delim(here('analyze', 'input', 'aor_demog_indicators.csv.gz'), delim='|') %>%
  arrange(aor, year) %>% 
  mutate(year = as.factor(year))

pc_scale = 100000

# Should get this from an external constants file along with other AOR characteristics
specific_area_of_responsibility <- arr %>%
  filter(aor == specific_aor) %>%
  distinct(area_of_responsibility) %>% 
  as.character()

```

This notebook provides a basic descriptive overview of ICE ERO-LESA enforcement data for the `r print(paste0(specific_area_of_responsibility, " (",  specific_aor, ")"))`.

# Total enforcement actions by FY

```{r fy_total, echo=FALSE, message=FALSE, include=TRUE}

enc_fy <- enc %>% 
  filter(aor == specific_aor) %>% 
  group_by(fy) %>% 
  summarize(n_encounters = n())

arr_fy <- arr %>% 
  filter(aor == specific_aor) %>% 
  group_by(fy) %>% 
  summarize(n_arrests = n())

rem_fy <- rem %>% 
  filter(aor == specific_aor) %>% 
  group_by(fy) %>% 
  summarize(n_removals = n())

dat_aor <- left_join(enc_fy, arr_fy, by='fy') %>% 
  left_join(rem_fy, by='fy')

p1 <- dat_aor %>%
  pivot_longer(cols=-c('fy')) %>% 
  ggplot(aes(x = as.factor(fy), y=value, color=name, group=name)) +
  geom_line() +
  labs(title = "Total ICE enforcement events per FY",
       subtitle = specific_area_of_responsibility) +
  xlab('') +
  ylab("Value") +
  ylim(0, NA)

ggplotly(p1)

```

The following chart displays basic comparative measures for the `r specific_area_of_responsibility`: difference between total encounters and total arrests (`diff_enc_arr`); and difference between total arrests and total removals (`diff_arr_rem`).

```{r aor_diff, echo=FALSE, message=FALSE, include=TRUE}
dat_diff <- dat_aor %>% 
  mutate(diff_enc_arr = n_encounters - n_arrests,
         diff_rem_arr = n_removals - n_arrests)

p2 <- dat_diff %>% 
  pivot_longer(cols=-c('fy')) %>% 
  filter(name %in% c("diff_enc_arr", "diff_rem_arr")) %>% 
  ggplot(aes(x = fy, y=value, color=name, group=name)) +
  geom_line() +
  # geom_hline(aes(yintercept=0)) +
  # annotate("text", x = 2, y = 0, label = "Fewer arrests\nMore arrests") +
  labs(title = "Difference between enforcement varibles",
       subtitle = specific_area_of_responsibility)

ggplotly(p2)

```

Here we compare trends for the `r specific_area_of_responsibility` with national trends:

```{r aor_v_national}

enc_fy_natl <- enc %>% 
  filter(
    event_date <= "2022-09-30") %>% 
  group_by(fy) %>% 
  summarize(n_encounters = n())

arr_fy_natl <- arr %>% 
  filter(
    arrest_date <= "2022-09-30") %>% 
  group_by(fy) %>% 
  summarize(n_arrests = n())

rem_fy_natl <- rem %>% 
  filter(
    departed_date <= "2022-09-30") %>% 
  group_by(fy) %>% 
  summarize(n_removals = n())

dat_natl <- left_join(enc_fy_natl, arr_fy_natl, by='fy') %>% 
  left_join(rem_fy_natl, by='fy')

dat_aor$group <- specific_area_of_responsibility
dat_natl$group <- "National total"

dat <- rbind(dat_aor, dat_natl)

p1 <- dat %>%
  rename(encounters = n_encounters,
         arrests = n_arrests,
         removals = n_removals) %>% 
  pivot_longer(cols=-c('fy', 'group')) %>% 
  ggplot(aes(x = as.numeric(as.character(fy)), y=value, color=name, group=name)) +
  geom_line() +
  labs(title = "Total ICE enforcement events, FY2012-22",
       subtitle = paste0("National total compared to ", specific_area_of_responsibility)) +
  xlab('Fiscal Year') +
  scale_x_continuous(breaks=seq(2012,2024,4),
                     minor_breaks=seq(2012, 2024, 1)) +
  ylab("") +
  ylim(0, NA) +
  scale_color_discrete(name = "Enforcement type") +
  facet_wrap(~group, scales = "free_y")

p1

```


## Percent change

Percent change in enforcement actions per FY compared to national.

```{r arrests_pct_chg, message=FALSE, warning=FALSE}

dat_pct <- dat %>% 
  group_by(group) %>% 
  mutate(across(starts_with('n_'), ~ (.x/lag(.x) - 1))) %>% 
  pivot_longer(cols = c(-fy, -group))

p3 <- dat_pct %>% 
  filter(!is.na(value)) %>% 
  ggplot() +
  geom_col(aes(x = fy, y = value, fill=name, group=name), position='dodge') +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(breaks=seq(2013,2024,4)) +
  facet_wrap(~group) +
  labs(title = "Annual % change total ICE arrests, FY12-22",
       subtitle = paste0(specific_area_of_responsibility, " versus national"))

p3

```


# Encounters

```{r encounters_pc_rank, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}

enc_per_aor <- enc %>% 
  filter(!is.na(aor),
         aor != "HQ",
        event_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n = n())

enc_pc_per_aor <- left_join(enc_per_aor, demog, by=c('fy' = 'year', 'aor' = 'aor')) %>% 
  group_by(aor) %>% 
  fill(contains('pop')) %>% 
  mutate(n_per_cap = (n / total_pop) * pc_scale,
         n_per_undocu = (n / undocu_pop) * pc_scale) %>% 
  arrange(fy, desc(n_per_cap)) %>% 
  group_by(fy) %>% 
  mutate(pc_rank = row_number())

pc_encounter_rank_fy12 <- as.numeric(enc_pc_per_aor[enc_pc_per_aor$aor == specific_aor & enc_pc_per_aor$fy == 2012, 'pc_rank'])
pc_encounter_rank_fy22 <- as.numeric(enc_pc_per_aor[enc_pc_per_aor$aor == specific_aor & enc_pc_per_aor$fy == 2022, 'pc_rank'])

p1 <- enc_pc_per_aor %>% 
  ggplot(aes(x = fy, y=n_per_cap, color=aor, group=aor)) +
  geom_line() +
  gghighlight(aor == specific_aor, use_direct_label = FALSE) +
  xlab("Fiscal Year") +
  ylab("Encounters per capita") +
  labs(title = "ICE encounters per 100,000 residents",
       subtitle = paste0(specific_area_of_responsibility, " highlighted"))

p1  

```

## Encounters by `gender`

```{r enc_gender, message=FALSE, warning=FALSE, include=TRUE}

p1 <- enc %>%
  filter(aor == specific_aor) %>% 
  count(fy, gender) %>% 
  ggplot(aes(x=fy, y=n, fill=gender)) +
  geom_col(position='fill') +
  scale_y_continuous(labels = scales::percent) +
  ylab("") +
  xlab('Fiscal Year') +
  labs(title="ICE encounters, % by gender",
       subtitle=specific_area_of_responsibility)

p1

```

## Encounters by `citizenship_country`

```{r cit, message=FALSE, warning=FALSE, include=TRUE}

cit <- enc %>%
  filter(aor == specific_aor) %>%
  mutate(citizenship_country = toupper(citizenship_country)) %>% 
  group_by(citizenship_country) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

p1 <- enc %>%
  filter(aor == specific_aor) %>%
  mutate(citizenship_country = case_when(
    citizenship_country %in% head(cit$citizenship_country, 15) ~ citizenship_country,
    TRUE ~ "ALL OTHERS"
  )) %>% 
  count(fy, citizenship_country) %>% 
  ggplot(aes(x=fy, y=n, fill=citizenship_country, color=citizenship_country)) +
  geom_col() +
  labs(title = paste0(specific_aor, " ICE encounters by country of citizenship (top 15)"))

ggplotly(p1)

```

## Encounters by `operation`

```{r enc_op_fy, message=FALSE, warning=FALSE, include=TRUE}

ops <- enc %>% 
  filter(aor == specific_aor) %>% 
  count(operation) %>% 
  arrange(desc(n))

p1 <- enc %>% 
  filter(aor == specific_aor) %>% 
  mutate(operation_short = case_when(operation %in% head(ops$operation, 10) ~ as.character(operation), 
                                         TRUE ~ "ALL OTHERS")) %>% 
  group_by(fy, operation_short) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n, fill=operation_short)) +
  geom_col() +
  labs(title = "Total ICE encounters per FY by operation")

ggplotly(p1)

```

## Encounters by `processing_disposition`

```{r enc_disp_fy, message=FALSE, warning=FALSE, include=TRUE}

aor_disps <- enc %>%
  filter(aor == specific_aor) %>% 
  count(processing_disposition) %>% 
  arrange(desc(n))

p1 <- enc %>% 
  mutate(disp_short = case_when(processing_disposition %in% head(aor_disps$processing_disposition, 10) ~ as.character(processing_disposition), 
                                         TRUE ~ "ALL OTHERS")) %>% 
  group_by(fy, disp_short) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = fy, y=n, fill=disp_short, color=disp_short)) +
  geom_col() +
  labs(title = "Total ICE encounters per FY by processing disposition")

ggplotly(p1, dynamicTicks = TRUE)

```

# Arrests

```{r arrests_pc_rank, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}

arr_per_aor <- arr %>% 
  filter(!is.na(aor),
         aor != "HQ",
        arrest_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n = n())

arr_pc_per_aor <- left_join(arr_per_aor, demog, by=c('fy' = 'year', 'aor' = 'aor')) %>% 
  group_by(aor) %>% 
  fill(contains('pop')) %>% 
  mutate(n_per_cap = (n / total_pop) * pc_scale,
         n_per_undocu = (n / undocu_pop) * pc_scale) %>% 
  arrange(fy, desc(n_per_cap)) %>% 
  group_by(fy) %>% 
  mutate(pc_rank = row_number())

pc_arrest_rank_fy12 <- as.numeric(arr_pc_per_aor[arr_pc_per_aor$aor == specific_aor & arr_pc_per_aor$fy == 2012, 'pc_rank'])
pc_arrest_rank_fy19 <- as.numeric(arr_pc_per_aor[arr_pc_per_aor$aor == specific_aor & arr_pc_per_aor$fy == 2019, 'pc_rank'])
pc_arrest_rank_fy22 <- as.numeric(arr_pc_per_aor[arr_pc_per_aor$aor == specific_aor & arr_pc_per_aor$fy == 2022, 'pc_rank'])

p1 <- arr_pc_per_aor %>% 
  ggplot(aes(x = as.factor(fy), y=n_per_cap, color=aor, group=aor)) +
  geom_line() +
  gghighlight(aor == specific_aor, use_direct_label = FALSE) +
  labs(title = "ICE arrests per 100,000 residents",
       subtitle = paste0(specific_area_of_responsibility, " highlighted"))

p1  

p2 <- arr_pc_per_aor %>% 
  ggplot(aes(x = as.factor(fy), y=pc_rank, color=aor, group=aor)) +
  geom_line() +
  gghighlight(aor == specific_aor, use_direct_label = FALSE) +
  scale_y_reverse() +
  labs(title = "ICE arrests per 100,000 residents, AOR rank",
       subtitle = paste0(specific_area_of_responsibility, " highlighted"))

p2

p3 <- arr_pc_per_aor %>% 
  filter(aor == specific_aor) %>% 
  ggplot(aes(x = as.factor(fy), y=n_per_cap, color=aor, group=aor)) +
  geom_line() +
  ylim(0, NA) +
  labs(title = "ICE arrests per 100,000 residents",
       subtitle = paste0(specific_area_of_responsibility))

p3  

```

```{r prop_enc_arr}

dat <- full_join(enc_per_aor, arr_per_aor, by = c("fy", "aor"), suffix = c(".enc", ".arr"))

dat <- dat %>% 
  mutate(diff_enc_arr = n.enc - n.arr,
         ratio_enc_arr = n.enc / n.arr)

p1 <- dat %>% 
  ggplot(aes(x = fy, y = diff_enc_arr, color = aor, group = aor)) +
  geom_line() +
  gghighlight(aor ==specific_aor)

p1

p2 <- dat %>% 
  ggplot(aes(x = fy, y = ratio_enc_arr, color = aor, group = aor)) +
  geom_line() +
  gghighlight(aor ==specific_aor)

p2

```

## Demographics
 
```{r gender, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}

p1 <- arr %>% 
  filter(aor == specific_aor,
         arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>%
  mutate(gender = toupper(gender)) %>% 
  count(fy, gender) %>% 
  ggplot(aes(x=fy, y=n, fill=gender)) +
  geom_col(position='fill') +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Total ICE arrests, % by gender",
       subtitle=specific_area_of_responsibility)

p1

```

```{r cit, echo=FALSE, message=FALSE, include=TRUE}

cit <- arr %>%
  filter(aor == specific_aor,
         arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>%
  mutate(citizenship_country = toupper(citizenship_country)) %>% 
  group_by(citizenship_country) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

cit_categories <- 10

p1 <- arr %>% 
  filter(aor == specific_aor,
        arrest_date <= "2022-09-30") %>%
  mutate(citizenship_country = case_when(
    citizenship_country %in% head(cit$citizenship_country, cit_categories) ~ citizenship_country,
    TRUE ~ "ALL OTHERS"
  )) %>% 
  count(fy, citizenship_country) %>% 
  ggplot(aes(x=fy, y=n, fill=citizenship_country)) +
  geom_col() +
  labs(title = paste0("Total ICE arrests by country of citizenship (top ", cit_categories, ")"),
       subtitle = specific_area_of_responsibility)

p1

```


```{r cit_rank, echo=FALSE, message=FALSE, include=TRUE}

cit_rank <- arr %>% 
  filter(aor == specific_aor,
         arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>%
  count(fy, citizenship_country) %>% 
  arrange(fy, desc(n), citizenship_country) %>% 
  group_by(fy) %>% 
  mutate(ranking = row_number())

p1 <- cit_rank %>% 
  filter(ranking <= 10) %>% 
  ggplot(aes(x = fy, y = ranking, color = citizenship_country, group = citizenship_country)) +
  geom_line(alpha = .7, size = 1) +
  geom_point(alpha = .7, size = 2) +
  scale_y_reverse() +
  labs(title = "Ranked country of citizenship for ICE arrests",
       subtitle = specific_area_of_responsibility)

ggplotly(p1)

```


```{r method, echo=FALSE, message=FALSE, include=TRUE}

methods <- arr %>% 
  filter(aor == specific_aor,
         arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>%
  count(arrest_method) %>% 
  arrange(desc(n))

top_methods <- methods %>%
  filter(n > 100)

arr <- arr %>% 
  mutate(arrest_method_short = case_when(arrest_method %in% unlist(top_methods$arrest_method) ~ as.character(arrest_method), 
                                         TRUE ~ "All others"))

p1 <- arr %>% 
  filter(aor == specific_aor,
         arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>%
  group_by(fy, arrest_method_short) %>%
  ggplot(aes(x = fy, fill=arrest_method_short)) +
  geom_bar(stat='count', position='stack')

ggplotly(p1)

dat <- arr %>% 
  filter(aor == specific_aor,
         arrest_method_short %in% c("CAP Local Incarceration",
                                    "CAP State Incarceration",
                                    "CAP Federal Incarceration"),
         arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>%
  count(arrest_method_short, fy)

p2 <- dat %>% 
  ggplot(aes(x = fy, y=n, fill=arrest_method_short)) +
  geom_col(position='dodge') +
  labs(title = "Total CAP arrests, FY12-22",
       subtitle = specific_area_of_responsibility)

p2

dat <- arr %>% 
  filter(arrest_method_short %in% c("CAP Local Incarceration",
                                    "CAP State Incarceration",
                                    "CAP Federal Incarceration"),
         arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>%
  count(arrest_method_short, fy)

p3 <- dat %>% 
  ggplot(aes(x = fy, y=n, fill=arrest_method_short)) +
  geom_col(position='dodge') +
  labs(title = "Total CAP arrests, FY12-22",
       subtitle = "National")

p3


```

```{r natl_vs_aor_methods}

dat1 <- arr %>% 
  filter(arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>% 
  count(fy, arrest_method_short) %>% 
  mutate(group = "National")

dat2 <- arr %>% 
  filter(aor == specific_aor,
         arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30") %>% 
  count(fy, arrest_method_short) %>%
  mutate(group = specific_area_of_responsibility)

dat <- rbind(dat1, dat2)

p1 <- dat %>% 
  filter(group == "National") %>% 
  ggplot(aes(x = fy, y = n, color = arrest_method_short, group=arrest_method_short)) +
  geom_line() +
  facet_wrap(~group, scales = "free_y")

p1

p2 <- dat %>% 
  filter(group == specific_area_of_responsibility) %>% 
  ggplot(aes(x = fy, y = n, color = arrest_method_short, group=arrest_method_short)) +
  geom_line() +
  facet_wrap(~group, scales = "free_y")

p2

```

# Removals

```{r remvoals_pc_rank, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}

rem_per_aor <- rem %>% 
  filter(!is.na(aor),
         aor != "HQ",
         departed_date >= "2011-10-01",
        departed_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n = n())

rem_pc_per_aor <- left_join(rem_per_aor, demog, by=c('fy' = 'year', 'aor' = 'aor')) %>% 
  group_by(aor) %>% 
  fill(contains('pop')) %>% 
  mutate(n_per_cap = (n / total_pop) * pc_scale,
         n_per_undocu = (n / undocu_pop) * pc_scale) %>% 
  arrange(fy, desc(n_per_cap)) %>% 
  group_by(fy) %>% 
  mutate(pc_rank = row_number())

pc_removals_rank_fy12 <- as.numeric(rem_pc_per_aor[rem_pc_per_aor$aor == specific_aor & rem_pc_per_aor$fy == 2012, 'pc_rank'])
pc_removals_rank_fy22 <- as.numeric(rem_pc_per_aor[rem_pc_per_aor$aor == specific_aor & rem_pc_per_aor$fy == 2022, 'pc_rank'])

p1 <- rem_pc_per_aor %>% 
  ggplot(aes(x = as.factor(fy), y=n_per_cap, color=aor, group=aor)) +
  geom_line() +
  gghighlight(aor == specific_aor) +
  labs(title = "ICE removals per 100,000 residents",
       subtitle = paste0(specific_area_of_responsibility, " highlighted"))

p1  

```
