---
title: "ICE ERO-LESA nationwide arrests data, FY12-22"
author: "UWCHR"
date: "2023-06-01"
output:
    html_document:
        html_preview: true
        toc: true
        toc_depth: 3
        toc_float: true
        code_folding: hide
---

# Data overview

This notebook presents a national overview of U.S. Immigration and Customs Enforcement (ICE) Enforcement and Removal Operations (ERO) Law Enforcement Systems and Analysis Division (LESA) data regarding nationwide ICE arrests obtained by the University of Washington Center for Human Rights pursuant to FOIA request 2022-ICFO-09023.

```{r setup, message=FALSE, warning=FALSE, include=TRUE}

options(scipen = 1000000)

library(pacman)

p_load(here, tidyverse, zoo, lubridate, ggplot2, plotly, gghighlight, viridis)

arr <- read_delim(here('analyze', 'input', 'ice_arrests_fy12-23ytd.csv.gz'), delim='|',
                  col_types = cols(aor = col_factor(),
                                  arrest_date = col_date(format="%m/%d/%Y"),
                                  departed_date = col_date(format="%m/%d/%Y"),
                                  apprehension_landmark = col_factor(),
                                  arrest_method = col_factor(),
                                  operation = col_factor(),
                                  processing_disposition = col_factor(),
                                  citizenship_country = col_factor(),
                                  gender = col_factor(),
                                  case_closed_date = col_date(format="%m/%d/%Y"),
                                  id = col_integer(),
                                  hashid = col_character()
                                  )) 

redacted <- c('removal_threat_level', 'apprehension_threat_level', 'alien_file_number')
redacted_text <- paste0('`', paste(unlist(redacted), collapse = '`, `'), '`')

arr <- arr %>% 
  dplyr::select(-all_of(redacted))

cy_months <- c("Jan","Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
fy_months <- c("Oct", "Nov", "Dec", "Jan","Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep")

arr <- arr %>% 
  mutate(aor = factor(aor, levels = sort(levels(arr$aor))),
         arrest_date = as_date(arrest_date, format="%m/%d/%Y"),
         year = year(arrest_date),
         month = factor(month(arrest_date, label=TRUE, abbr=TRUE), levels = fy_months),
         year_mth = zoo::as.yearmon(arrest_date),
         fy = as.factor(substr(quarter(arrest_date, fiscal_start=10, type="year.quarter"), 1,4)),
         gender = toupper(gender),
         operation = toupper(operation),
         processing_disposition = toupper(processing_disposition),
         citizenship_country = factor(toupper(citizenship_country)),
         apprehension_landmark = toupper(str_squish(apprehension_landmark))) %>% 
  filter(arrest_date >= "2011-10-01",
         arrest_date <= "2022-09-30")

```

An arrest occurs when an individual is taken into custody by ICE and removal proceedings initiated against them.[^1]

The arrests dataset (`arr`) includes `r nrow(arr)` observations of `r length(arr)` variables; `r length(redacted)` fully redacted fields (`r redacted_text`) are dropped from analysis.

The following provides an summary of dataset characteristics via `skimr::skim(arr)`:

```{r skim, message=FALSE, warning=FALSE, include=TRUE}

skimr::skim(arr)

```

## Field definitions

#### Original dataset fields
- `aor`: ICE Area of Responsibility associated with arrest
- `arrest_date`: Date of arrest
- `departed_date`: Date of departure
- `case_closed_date`: Date of closure of case
- `arrest_method`: ICE ERO division or category associated with arrest
- `apprehension_landmark`: Landmark or entity associated with arrest
- `operation`: Operation associated with arrest
- `processing_disposition`: Status of removal proceedings associated with event
- `citizenship_country`: Country of citizenship of arrested individual
- `gender`: Gender of arrested individual
- `apprehension_threat_level`: Fully redacted in original dataset
- `removal`_threat_level`: Fully redacted in original dataset
- `alien_file_number`: Unique individual identifier for arrested individual, fully redacted in original dataset

#### Additional fields created by UWCHR
- `id`: Sequential record identifier (not individual identifier)
- `hashid`: Unique record hash (not individual identifier)
- `year`: Calendar year derived from `arrest_date`
- `month`: Abbreviated month derived from `arrest_date`
- `year_mth`: Calendar year and month derived from `arrest_date`
- `fy`: U.S. government fiscal year (Oct.-Sept.) derived from `arrest_date`

# Total arrests

```{r fy_total, message=FALSE, include=TRUE}

p1 <- arr %>% 
  group_by(fy) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n)) +
  geom_col() +
  labs(title = "Total ICE arrests per FY")

p1

p2 <- arr %>%
  group_by(year_mth) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = year_mth, y = n)) +
  geom_line(aes(group=1)) +
  ylim(0, NA) +
  labs(title = "Total nationwide ICE arrests per month")

p2

p3 <- arr %>%
  group_by(fy, month) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = month, y = n, color = fy, group = fy)) +
  geom_line() +
  ylim(0, NA) +
  scale_color_viridis_d() +
  labs(title = "Total nationwide ICE arrests per month")

p3

```

# Basic demographics

## Gender

Increasing proportion of females arrested since FY 2021:
 
```{r gender, message=FALSE, warning=FALSE, include=TRUE}

# arr %>%
#   mutate(gender = toupper(gender)) %>% 
#   group_by(gender) %>% 
#   summarize(n = n())

p1 <- arr %>% 
  count(fy, gender) %>% 
  ggplot(aes(x=fy, y=n, fill=gender)) +
  geom_col(position='fill') +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Total ICE arrests, % by gender")

p1

```

## Country of citizenship

Changing composition of arrest nationality: Mexico, Guatemala, El Salvador decrease; increase in Venezeula, Colombia, Nicaragua.

```{r cit, message=FALSE, include=TRUE}

cit <- arr %>%
  mutate(citizenship_country = toupper(citizenship_country)) %>% 
  group_by(citizenship_country) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

p1 <- arr %>% 
  mutate(citizenship_country = case_when(
    citizenship_country %in% head(cit$citizenship_country, 15) ~ citizenship_country,
    TRUE ~ "ALL OTHERS"
  )) %>% 
  count(fy, citizenship_country) %>% 
  ggplot(aes(x=fy, y=n, fill=citizenship_country, color=citizenship_country)) +
  geom_col() +
  labs(title = "Total ICE arrests by country of citizenship (top 15)")

ggplotly(p1)

```

```{r cit_rank, message=FALSE, include=TRUE}

cit_rank <- arr %>% 
  count(fy, citizenship_country) %>% 
  arrange(fy, desc(n), citizenship_country) %>% 
  group_by(fy) %>% 
  mutate(ranking = row_number())

p1 <- cit_rank %>% 
  filter(ranking <= 10) %>% 
  ggplot(aes(x = fy, y = ranking, group = citizenship_country)) +
  geom_line(aes(color = citizenship_country), size = 1) +
  geom_point(aes(color = citizenship_country), size = 2) +
  scale_y_reverse(breaks = seq(1,10)) +
  labs(title = "Ranked country of citizenship for ICE arrests")

ggplotly(p1)

# How to illustrate proportional increase of small groups, e.g. ANGOLA? Percent change by citizenship group?

```

# Total arrests by AOR

Below is an interactive chart of total ICE arrests per FY by AOR:

```{r annual, message=FALSE, include=TRUE}

p1 <- arr %>% 
  group_by(fy, aor) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n, color=aor, group=aor)) +
  geom_line() +
  labs(title = "Total ICE arrests per FY by AOR")

ggplotly(p1)

```

## Percent change

Percent change in arrests per FY nationally and by AOR. Note comparison between groups may be misleading because smaller population AORs will be more volatile.

```{r arrests_pct_chg, message=FALSE, warning=FALSE}

natl_pct_chg <- arr %>%
  group_by(fy) %>%
  summarize(n = n()) %>% 
  mutate(pct_change = (n/lag(n) - 1))

p1 <- natl_pct_chg %>% 
  ggplot(aes(x = fy, y = pct_change)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(title="FY % change in total ICE arrests")

p1

aor_pct_chg <- arr %>%
  filter(!is.na(aor)) %>% 
  group_by(fy, aor) %>%
  summarize(n = n()) %>% 
  group_by(aor) %>% 
  arrange(fy, .by_group=TRUE) %>% 
  mutate(pct_change = (n/lag(n) - 1))

p2 <- aor_pct_chg %>% 
  ggplot(aes(x = fy, y = pct_change)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(breaks=seq(2012, 2022, 2)) +
  facet_wrap(~aor)  +
  labs(title="FY % change in total ICE arrests per AOR") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p2

```

# Arrests by `arrest_method`

```{r method_setup, message=FALSE, include=TRUE}

methods <- arr %>% 
  count(arrest_method) %>% 
  arrange(desc(n))

top_methods <- methods %>% 
  filter(n > 5000)

arr <- arr %>% 
  mutate(arrest_method_short = case_when(arrest_method %in% unlist(top_methods$arrest_method) ~ as.character(arrest_method), 
                                         TRUE ~ "All others"))

```

```{r national_arrest_method, message=FALSE, warning=FALSE}

p1 <- arr %>% 
  group_by(fy, arrest_method_short) %>%
  ggplot(aes(x = fy, fill=arrest_method_short)) +
  geom_bar(stat='count', position='stack')

ggplotly(p1)

```

```{r arr_method_pct_chg, message=FALSE, warning=FALSE}

method_pct_chg <- arr %>%
  filter(!is.na(aor)) %>% 
  group_by(fy, arrest_method_short) %>%
  summarize(n = n()) %>% 
  group_by(arrest_method_short) %>% 
  arrange(fy, .by_group=TRUE) %>% 
  mutate(pct_change = (n/lag(n) - 1))

p1 <- method_pct_chg %>% 
  ggplot(aes(x = fy, y = pct_change)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(breaks=seq(2012, 2022, 2)) +
  facet_wrap(~arrest_method_short, scales='free_y')  +
  labs(title="FY % change in total ICE arrests by arrest method") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p1

```

```{r method_per_aor, message=FALSE, warning=FALSE}

p2 <- arr %>% 
  filter(!is.na(aor),
         aor != "HQ") %>% 
  group_by(aor, fy, arrest_method_short) %>%
  ggplot(aes(x = fy, fill=arrest_method_short)) +
  geom_bar(stat='count') +
  scale_x_discrete(breaks=seq(2012, 2022, 2)) +
  facet_wrap(~aor) +
  labs(title="Total ICE arrests by arrest method per AOR") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=1))

ggplotly(p2)

p3 <- arr %>% 
  filter(!is.na(aor),
         aor != "HQ",
         arrest_method_short == "ERO Reprocessed Arrest") %>% 
  group_by(aor, fy, arrest_method_short) %>%
  ggplot(aes(x = fy, fill=arrest_method_short)) +
  geom_bar(stat='count') +
  scale_x_discrete(breaks=seq(2012, 2022, 2)) +
  facet_wrap(~aor) +
  labs(title="Total ICE arrests by arrest method per AOR") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=1))

p3

```

```{r arrest_method_sex}

p1 <- arr %>% 
  count(fy, gender, arrest_method_short) %>% 
  ggplot(aes(x=fy, y=n, fill=gender)) +
  geom_col(position='fill') +
  facet_wrap(~arrest_method_short) +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Total ICE arrests, % by gender")

p1

```

# Processing disposition

```{r arr_disp, message=FALSE, warning=FALSE, include=TRUE}

disps <- arr %>% 
  count(processing_disposition) %>% 
  arrange(desc(n))

top_disps <- disps %>% 
  filter(n > 10000)

arr <- arr %>% 
  mutate(disp_short = case_when(processing_disposition %in% unlist(top_disps$processing_disposition) ~ as.character(processing_disposition), 
                                         TRUE ~ "ALL OTHERS"))

p1 <- arr %>%
  group_by(fy, disp_short) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n, fill=disp_short)) +
  geom_col() +
  labs(title = "Total ICE arrests per FY by processing disposition")

ggplotly(p1)

```

# Apprehension Landmark

Overview of most common arrest `apprehension_landmark` values gives a sense of the diversity of this category, which includes `r length(unique(arr$apprehension_landmark))` unique values; note inclusion of general values likely denoting the ICE sub-office or divison responsible for the arrest rather than a precise location. Closer inspection is recommended at the AOR level.

```{r arr_landmark, message=FALSE, warning=FALSE, include=TRUE}

landmarks_per_aor <- arr %>%
  group_by(aor) %>%
  summarize(n = n_distinct(apprehension_landmark))

landmarks <- arr %>% 
  count(apprehension_landmark) %>% 
  arrange(desc(n))

p1 <- arr %>% 
  mutate(apprehension_landmark = case_when(apprehension_landmark %in% head(landmarks$apprehension_landmark, 15) ~ as.character(apprehension_landmark), 
                                         TRUE ~ "ALL OTHERS")) %>% 
  group_by(fy, apprehension_landmark) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n, fill=apprehension_landmark)) +
  geom_col() +
  labs(title = "Total ICE arrests per FY by `apprehension_landmark` (top 15)")

ggplotly(p1)

```

[^1]: For discussion of ICE's definition of "arrests", see American Immigration Council, "Changing Patterns of Interior Immigration Enforcement in the United States, 2016 - 2018", July 2019: https://www.americanimmigrationcouncil.org/research/interior-immigration-enforcement-united-states-2016-2018
