---
title: "ICE ERO-LESA enforcement actions data, FY12-22"
author: "UWCHR"
date: "2024-05-31"
output:
    html_document:
        html_preview: true
        toc: true
        toc_depth: 3
        toc_float: true
        code_folding: hide
---

```{r setup, message=FALSE, warning=FALSE, include=TRUE}

options(scipen = 1000000)

library(pacman)

p_load(here, tidyverse, zoo, lubridate, ggplot2, plotly, gghighlight, sf, tigris, ggrepel, ggmap, tidygeocoder, viridis)

```

# Data overview

This notebook presents a national overview of U.S. Immigration and Customs Enforcement (ICE) Enforcement and Removal Operations (ERO) Law Enforcement Systems and Analysis Division (LESA) data regarding nationwide encounters, arrests, and removals obtained by the University of Washington Center for Human Rights (UWCHR) pursuant to FOIA request 2022-ICFO-09023.

## Datasets

In each dataset, a record relates to an enforcement event, not an individual. Individuals may be involved in multiple enforcement events of each category. Original datasets do not include unique individual identifiers; it is not possible to link a given encounter to an associated arrest, or an arrest to an associated removal.[^1]

Datasets have been minimally cleaned and standardized; in the following sections we load each dataset and perform some basic additional standardization and creation of convenience variables for various units of time, etc.

For additional detail, see respective descriptive notebooks for each enforcement dataset:

- [Encounters](encounters.html)
- [Arrests](arrests.html)
- [Removals](removals.html)

### Encounters

```{r load_encounters, message=FALSE, warning=FALSE, include=TRUE}

# ENCOUNTERS DATA

enc <- read_delim(here('analyze', 'input', 'ice_encounters_fy12-23ytd.csv.gz'), delim='|',
                  col_types = cols(aor = col_factor(),
                                   event_date = col_character(),
                                   landmark = col_character(),
                                   operation = col_factor(),
                                   processing_disposition = col_factor(),
                                   citizenship_country = col_factor(),
                                   gender = col_factor(),
                                   hashid = col_character(),
                                   id = col_number()),
                  show_col_types = FALSE)

redacted <- c('encounter_threat_level', 'alien_file_number')
redacted_text <- paste0('`', paste(unlist(redacted), collapse = '`, `'), '`')

enc <- enc %>% 
  dplyr::select(-redacted)

enc <- enc %>% 
  mutate(aor = factor(aor, levels = sort(levels(enc$aor))),
         event_date = as_date(event_date, format="%m/%d/%Y"),
         year = year(event_date),
         month = month(event_date, label=TRUE, abbr=TRUE),
         year_mth = zoo::as.yearmon(event_date),
         fy_quarter = as.factor(quarter(event_date, fiscal_start=10, type="year.quarter")),
         fy = as.factor(substr(fy_quarter, 1,4)),
         gender = toupper(gender),
         operation = toupper(operation),
         processing_disposition = toupper(processing_disposition),
         citizenship_country = toupper(citizenship_country))

```

An encounter occurs when an individual is subjected to revision of admissibility/removability by ICE, and may or may not lead to an arrest.[^2]

The encounters dataset (`enc`) includes `r nrow(enc)` observations of `r length(enc)` variables; `r length(redacted)` fully redacted fields (`r redacted_text`) are dropped from analysis. For additional detail, see the [Encounters](encounters.html) notebook.

```{r glimpse_encounters, message=FALSE, warning=FALSE, include=TRUE}

glimpse(enc)

```

### Arrests

```{r load_arrests, message=FALSE, warning=FALSE, include=TRUE}

# ARRESTS DATA

arr <- read_delim(here('analyze', 'input', 'ice_arrests_fy12-23ytd.csv.gz'), delim='|',
                  col_types = cols(aor = col_factor(),
                                  arrest_date = col_date(format="%m/%d/%Y"),
                                  departed_date = col_date(format="%m/%d/%Y"),
                                  apprehension_landmark = col_factor(),
                                  arrest_method = col_factor(),
                                  operation = col_factor(),
                                  processing_disposition = col_factor(),
                                  citizenship_country = col_factor(),
                                  gender = col_factor(),
                                  case_closed_date = col_date(format="%m/%d/%Y"),
                                  id = col_integer(),
                                  hashid = col_character()
                                  )) 

redacted <- c('removal_threat_level', 'apprehension_threat_level', 'alien_file_number')
redacted_text <- paste0('`', paste(unlist(redacted), collapse = '`, `'), '`')

arr <- arr %>% 
  dplyr::select(-all_of(redacted))

arr <- arr %>% 
  mutate(aor = factor(aor, levels = sort(levels(arr$aor))),
         arrest_date = as_date(arrest_date, format="%m/%d/%Y"),
         year = year(arrest_date),
         month = month(arrest_date, label=TRUE, abbr=TRUE),
         year_mth = zoo::as.yearmon(arrest_date),
         fy_quarter = as.factor(quarter(arrest_date, fiscal_start=10, type="year.quarter")),
         fy = as.factor(substr(fy_quarter, 1,4)),
         citizenship_country = as.factor(toupper(citizenship_country)))

```

An arrest occurs when an individual is taken into custody by ICE and removal proceedings initiated against them.

The arrests dataset (`arr`) includes `r nrow(arr)` observations of `r length(arr)` variables; `r length(redacted)` fully redacted fields (`r redacted_text`) are dropped from analysis. For additional detail, see the [Arrests](arrests.html) notebook.

```{r glimpse_arrests, message=FALSE, warning=FALSE, include=TRUE}

glimpse(arr)

```

### Removals

```{r load_removals, message=FALSE, warning=FALSE, include=TRUE}

# REMOVALS DATA

pd_dict <- read_delim(here('share', 'hand', 'processing_disp.csv'), delim='|')

rem <- read_delim(here('analyze', 'input', 'ice_removals_fy12-23ytd.csv.gz'), delim='|',
                  col_types = cols(aor = col_character(),
                                   arrest_date = col_date(format="%m/%d/%Y"),
                                  departed_date = col_date(format="%m/%d/%Y"),
                                  case_close_date = col_date(format="%m/%d/%Y"),
                                  removal_date = col_date(format="%m/%d/%Y"),
                                  apprehension_method_code = col_character(),
                                  processing_disposition_code = col_factor(),
                                  citizenship_country = col_factor(),
                                  gender = col_factor(),
                                  final_charge_section = col_factor(),
                                  id = col_integer(),
                                  hashid = col_character()
                                  )) 

# glimpse(rem)

redacted <- c('removal_threat_level', 'alien_file_number')
redacted_text <- paste0('`', paste(unlist(redacted), collapse = '`, `'), '`')

rem <- rem %>% 
  dplyr::select(-redacted, -case_closed_date)

rem <- rem %>% 
  mutate(aor = factor(aor, levels = sort(levels(rem$aor))),
         year = year(departed_date),
         month = month(departed_date, label=TRUE, abbr=TRUE),
         year_mth = zoo::as.yearmon(departed_date),
         processing_disp = toupper(coalesce(processing_disposition_code, processing_disposition)),
         fy_quarter = as.factor(quarter(departed_date, fiscal_start=10, type="year.quarter")),
         fy = as.factor(substr(fy_quarter, 1,4)))

rem <- left_join(rem, pd_dict, by=c('processing_disp' = 'processing_disposition_raw'))

```

A removal occurs when an individual is issued a final order of removal and departs the United States via deportation or voluntary return.

The removals dataset (`rem`) includes `r nrow(rem)` observations of `r length(rem)` variables; `r length(redacted)` fully redacted fields (`r redacted_text`) are dropped from analysis. For additional detail, see the [Removals](removals.html) notebook.

```{r glimpse_removals, message=FALSE, warning=FALSE, include=TRUE}

glimpse(rem)

```

```{r load_supplemental, message=FALSE, warning=FALSE, include=TRUE}

# SUPPLEMENTAL DATA

# AOR geographic boundaries by county
county_aor <- read_delim(here('share', 'hand', 'county_aor.csv'), delim = ',')
aor_field_office <- read_delim(here('share', 'hand', 'aor_ero_field_office.csv'), delim = ',')

# Select AOR-level demographics obtained via `tidycensus`
# We currently lack demographics for years following 2019, so we use 2019 values for subsequent years
# TO DO: Update demographics for more recent years
demog <- read_delim(here('analyze', 'input', 'aor_demog_indicators.csv.gz'), delim='|') %>%
  arrange(aor, year) %>% 
  mutate(year = as.character(year))

# Google Maps API key required to render maps, see `??ggmap::register_google`
ggkey = Sys.getenv("GOOGLEGEOCODE_API_KEY")
register_google(key = ggkey)

# Scale for per capita calculations
pc_scale = 100000

```

# Basic descriptive analysis

The following sections provide a basic descriptive overview of the three enforcement datasets.

## Total enforcement actions per FY

All three enforcement metrics peaked nationwide in FY 2012; we lack data for prior years. Note total arrests exceed total removals for first time in FY 2021; significant increase in rate of encounters to arrests in FY 2022.

```{r fy_total, message=FALSE, include=TRUE}

enc_fy <- enc %>% 
  filter(event_date <= "2022-09-30") %>% 
  group_by(fy) %>% 
  summarize(n_encounters = n())

arr_fy <- arr %>% 
  filter(arrest_date <= "2022-09-30") %>% 
  group_by(fy) %>% 
  summarize(n_arrests = n())

rem_fy <- rem %>% 
  filter(departed_date <= "2022-09-30") %>% 
  group_by(fy) %>% 
  summarize(n_removals = n())

dat <- left_join(enc_fy, arr_fy, by='fy') %>% 
  left_join(rem_fy, by='fy')

p1 <- dat %>%
  pivot_longer(cols=-c('fy')) %>% 
  ggplot(aes(x = as.factor(fy), y=value, color=name, group=name)) +
  geom_line() +
  labs(title = "Total ICE enforcement events per FY") +
  xlab('') +
  ylab("Value") +
  ylim(0, NA)

ggplotly(p1)

```

## ICE Areas of Responsibility (AOR)

ICE ERO divides the country into 24 Areas of Responsibility (AORs) and Field Offices, mapped below. This is the only common level of geographic organization shared between the three enforcement datasets.

Most AORs correspond to one or more U.S. states and territories; the states of CA, NY, and TX are divided between two or more AORs.[^3]

For additional detail, see the following descriptive notebooks for each AOR:

[ATL](aor/ATL.html) - [BAL](aor/BAL.html) - [BOS](aor/BOS.html) - [BUF](aor/BUF.html) - [CHI](aor/CHI.html) - [DAL](aor/DAL.html) - [DEN](aor/DEN.html) - [DET](aor/DET.html) - [ELP](aor/ELP.html) - [HOU](aor/HOU.html) - [LOS](aor/LOS.html) - [MIA](aor/MIA.html) - [NEW](aor/NEW.html) - [NOL](aor/NOL.html) - [NYC](aor/NYC.html) - [PHI](aor/PHI.html) - [PHO](aor/PHO.html) - [SEA](aor/SEA.html) - [SFR](aor/SFR.html) - [SLC](aor/SLC.html) - [SNA](aor/SNA.html) - [SND](aor/SND.html) - [SPM](aor/SPM.html) - [WAS](aor/WAS.html)

```{r maps_setup, message=FALSE, include=FALSE}
  
states <- unique(county_aor$geoid_state)

all_counties <- counties(state=states) %>%
  shift_geometry()

# Combine county-level geometries into an AOR geometry

aor_geom <- left_join(all_counties, county_aor, by=c('GEOID' = "geoid")) %>%
  group_by(aor) %>%
  summarize(geometry = sf::st_union(geometry))

aor_field_office <- read_delim(here('share', 'hand', 'aor_ero_field_office.csv'), delim = ',')

aor_field_office_sf <- aor_field_office %>%
  st_as_sf(coords = c("long", "lat"), crs=4326)

```

```{r map_aor, message=FALSE, include=TRUE}

p1 <- aor_geom %>%
  ggplot(aes(fill=aor)) +
  geom_sf(color='black') +
  geom_sf(data = aor_field_office_sf, aes(geometry = geometry), show.legend = FALSE) +
  ggrepel::geom_label_repel(
    data = aor_field_office_sf,
    aes(label = aor, geometry = geometry),
    fill = alpha(c("white"),0.5),
    color = "black",
    stat = "sf_coordinates",
    min.segment.length = 0,
    force=2,
    show.legend = FALSE
  ) +
  theme_void() +
  labs(title = "ICE AORs and Field Office locations")

p1

```

## Enforcement by AOR

Each enforcement dataset categorizes enforcement actions at the level of ICE Areas of Responsibility (AORs). Here we provide an overview of total enforcement actions per Fiscal Year across all AORs. Note wide distribution in rates of enforcement actions among AORs and over time.

```{r enforce_by_aor, message=FALSE, include=TRUE}

enc_fy_aor <- enc %>% 
  filter(event_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n_encounters = n()) %>% 
  mutate(fy = as.numeric(as.character(fy))) %>% 
  ungroup()

arr_fy_aor <- arr %>% 
  filter(arrest_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n_arrests = n()) %>% 
  mutate(fy = as.numeric(as.character(fy))) %>% 
  ungroup()

rem_fy_aor <- rem %>% 
  filter(departed_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n_removals = n()) %>% 
  mutate(fy = as.numeric(as.character(fy))) %>% 
  ungroup()

dat <- left_join(enc_fy_aor, arr_fy_aor, by=c('fy', 'aor')) %>% 
  left_join(rem_fy_aor, by=c('fy', 'aor'))

dat <- dat %>% 
  filter(!is.na(aor)) %>% 
  mutate(across(everything(), ~replace_na(., 0))) %>% 
  mutate(diff_enc_arr = n_encounters - n_arrests,
         diff_rem_arr = n_removals - n_arrests)
  

p1 <- dat %>%
  filter(!is.na(aor)) %>% 
  dplyr::select(-contains('diff')) %>% 
  pivot_longer(cols=-c('fy', 'aor')) %>% 
  ggplot(aes(x = as.factor(fy), y=value, color=name, group=name)) +
  geom_line() +
  # scale_y_log10() +
  scale_x_discrete(breaks=seq(2013,2024,4)) +
  labs(title = "Total ICE enforcement events per FY") +
  facet_wrap(~aor) +
  # facet_wrap(~aor, scales='free_y') +
  xlab('') +
  ylab("Value")

p1

```

Two simple comparative measures are the annual differences between total encounters and arrests; and between total removals and arrests per AOR.

```{r enforce_diff}

p2 <- dat %>%
  filter(!is.na(aor)) %>% 
  dplyr::select(fy, aor, contains('diff')) %>% 
  pivot_longer(cols=-c('fy', 'aor')) %>% 
  ggplot(aes(x = as.factor(fy), y=value, color=name, group=name)) +
  geom_line() +
  scale_x_discrete(breaks=seq(2013,2024,4)) +
  labs(title = "Difference between ICE enforcement events per FY") +
  facet_wrap(~aor) +
  xlab('') +
  ylab("Value")

p2

# b1 <- dat %>%
#   filter(!is.na(aor)) %>% 
#   dplyr::select(fy, aor, contains('diff')) %>% 
#   pivot_longer(cols=-c('fy', 'aor')) %>% 
#   ggplot(aes(x = as.factor(name), y=value, color=name, group=name)) +
#   geom_boxplot()
# 
# b1

```

## Rates of enforcement per capita

```{r encounters_pc, message=FALSE, warning=FALSE, include=TRUE}

enc_per_aor <- enc %>% 
  filter(!is.na(aor),
         aor != "HQ",
         event_date >= "2011-10-01",
        event_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n = n())

enc_pc_per_aor <- left_join(enc_per_aor, demog, by=c('fy' = 'year', 'aor' = 'aor')) %>% 
  group_by(aor) %>% 
  fill(contains('pop')) %>% 
  mutate(n_per_cap = (n / total_pop) * pc_scale,
         n_per_undocu = (n / undocu_pop) * pc_scale) %>% 
  arrange(fy, desc(n_per_cap)) %>% 
  group_by(fy) %>% 
  mutate(pc_rank = row_number())

p1 <- enc_pc_per_aor %>% 
  ggplot(aes(x = as.factor(fy), y=n_per_cap, color=aor, group=aor)) +
  geom_line() +
  labs(title = "ICE encounters per 100,000 residents")

ggplotly(p1)  

```

```{r arrests_pc, message=FALSE, warning=FALSE, include=TRUE}

arr_per_aor <- arr %>% 
  filter(!is.na(aor),
         aor != "HQ",
         arrest_date >= "2011-10-01",
        arrest_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n = n())

arr_pc_per_aor <- left_join(arr_per_aor, demog, by=c('fy' = 'year', 'aor' = 'aor')) %>% 
  group_by(aor) %>% 
  fill(contains('pop')) %>% 
  mutate(n_per_cap = (n / total_pop) * pc_scale,
         n_per_undocu = (n / undocu_pop) * pc_scale) %>% 
  arrange(fy, desc(n_per_cap)) %>% 
  group_by(fy) %>% 
  mutate(pc_rank = row_number())

p1 <- arr_pc_per_aor %>% 
  ggplot(aes(x = as.factor(fy), y=n_per_cap, color=aor, group=aor)) +
  geom_line() +
  labs(title = "ICE arrests per 100,000 residents")

ggplotly(p1)  

```

```{r removals_pc, message=FALSE, warning=FALSE, include=TRUE}

rem_per_aor <- rem %>% 
  filter(!is.na(aor),
         aor != "HQ",
         departed_date >= "2011-10-01",
         departed_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n = n())

rem_pc_per_aor <- left_join(rem_per_aor, demog, by=c('fy' = 'year', 'aor' = 'aor')) %>% 
  group_by(aor) %>% 
  fill(contains('pop')) %>% 
  mutate(n_per_cap = (n / total_pop) * pc_scale,
         n_per_undocu = (n / undocu_pop) * pc_scale) %>% 
  arrange(fy, desc(n_per_cap)) %>% 
  group_by(fy) %>% 
  mutate(pc_rank = row_number())

p1 <- rem_pc_per_aor %>% 
  ggplot(aes(x = as.factor(fy), y=n_per_cap, color=aor, group=aor)) +
  geom_line() +
  labs(title = "ICE removals per 100,000 residents")

ggplotly(p1)  

```

## Mapping enforcement

We can map enforcement variables at the AOR level as below:

```{r enforcement_per_capita}

# Note here we use 2019 demographic data for subsequent
dat_map <- dat %>%
  mutate(fy = as.character(fy)) %>%
  full_join(demog, by=c('aor' = 'aor', 'fy' = 'year')) %>%
  group_by(aor) %>%
  fill(contains('pop')) %>%
  left_join(aor_geom, by='aor')

p0 <- dat_map %>%
  filter(fy == 2022,
         aor != "HQ") %>%
  mutate(encounters_pc = n_encounters / total_pop * pc_scale) %>%
  ggplot(aes(geometry=geometry, fill=encounters_pc)) +
  geom_sf() +
  scale_fill_viridis_c() +
  theme_void() +
  labs(title = "ICE encounters per capita, FY 2022",
       caption = "Per est. 100,000 pop., 2019 (tidycensus)")

p1 <- dat_map %>%
  filter(fy == 2022,
         aor != "HQ") %>%
  mutate(arrests_pc = n_arrests / total_pop * pc_scale) %>%
  ggplot(aes(geometry=geometry, fill=arrests_pc)) +
  geom_sf() +
  scale_fill_viridis_c() +
  theme_void() +
  labs(title = "ICE arrests per capita, FY 2022",
       caption = "Per est. 100,000 pop., 2019 (tidycensus)")

p2 <- dat_map %>%
  filter(fy == 2022,
         aor != "HQ") %>%
  mutate(removals_pc = n_removals / total_pop * pc_scale) %>%
  ggplot(aes(geometry=geometry, fill=removals_pc)) +
  geom_sf() +
  scale_fill_viridis_c() +
  theme_void() +
  labs(title = "ICE removals per capita, FY 2022",
       caption = "Per est. 100,000 pop., 2019 (tidycensus)")

p0
p1
p2

```

[^1]: Unique values have been assigned for each record: `id` values denote the sequence of records in the original concatenated datasets; `hashid` values are unique identifiers for each record and are included for validation purposes (e.g. validating join operations, checking whether data has been transformed following assignment of `hashid` values, etc.).]

[^2]: For discussion of ICE's definitions of "encounters", "arrests", and "removals", see American Immigration Council, "Changing Patterns of Interior Immigration Enforcement in the United States, 2016 - 2018", July 2019: https://www.americanimmigrationcouncil.org/research/interior-immigration-enforcement-united-states-2016-2018

[^3]: An additional "HQ" AOR does not have an associated territorial jurisdiction. Some ICE datasets also refer to a "HAL" AOR for the south Texas region centering on Harlingen, TX; this AOR overlaps with the HOU and SNA regions as displayed in this notebook.
