---
title: "FY12-23YTD ICE ERO removals descriptive analysis"
author: "UWCHR"
date: "2023-03-27"
output:
    html_document:
        html_preview: true
        toc: true
        toc_depth: 3
        toc_float: true
        code_folding: hide
---
 

```{r setup, echo=FALSE, message=FALSE, include=TRUE}

library(pacman)

p_load(here, tidyverse, zoo, lubridate, ggplot2, plotly, gghighlight)

pd_dict <- read_delim(here('share', 'hand', 'processing_disp.csv'), delim='|')

rem <- read_delim(here('us-fy12-23', 'analyze', 'input', 'removals.csv.gz'), delim='|',
                  col_types = cols(aor = col_character(),
                                   arrest_date = col_date(format="%m/%d/%Y"),
                                  departed_date = col_date(format="%m/%d/%Y"),
                                  case_close_date = col_date(format="%m/%d/%Y"),
                                  removal_date = col_date(format="%m/%d/%Y"),
                                  apprehension_method_code = col_character(),
                                  processing_disposition_code = col_factor(),
                                  citizenship_country = col_factor(),
                                  gender = col_factor(),
                                  final_charge_section = col_factor(),
                                  id = col_integer(),
                                  hashid = col_character()
                                  )) 

glimpse(rem)

redacted <- c('removal_threat_level', 'alien_file_number')

rem <- rem %>% 
  dplyr::select(-redacted, -case_closed_date)

rem <- rem %>% 
  mutate(year = year(departed_date),
         month = month(departed_date, label=TRUE, abbr=TRUE),
         year_mth = zoo::as.yearmon(departed_date),
         processing_disp = toupper(coalesce(processing_disposition_code, processing_disposition)),
         fy =substr(quarter(departed_date, fiscal_start=10, type="year.quarter"), 1,4))

rem <- left_join(rem, pd_dict, by=c('processing_disp' = 'processing_disposition_raw'))

# skimr::skim(rem)

```

# Total removals per FY

Major decrease in removals by ICE, but note CBP expulsions at Southern border are not counted here.

```{r fy_total, echo=FALSE, message=FALSE, include=TRUE}

p1 <- rem %>% 
  filter(departed_date >= "2011-10-01",
         departed_date <= "2022-09-30"
         ) %>% 
  group_by(fy) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n)) +
  geom_col() +
  labs(title = "Total removals per FY")

p1

```

# Basic demographics

```{r gender, echo=FALSE, message=FALSE, include=TRUE}

# rem %>%
#   mutate(gender = tolower(gender)) %>% 
#   group_by(gender) %>% 
#   summarize(n = n())

p1 <- rem %>% 
  filter(departed_date >= "2011-10-01",
       departed_date <= "2022-09-30") %>% 
  count(fy, gender) %>% 
  ggplot(aes(x=fy, y=n, fill=gender)) +
  geom_col(position='fill') +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Total ICE removals, % by gender")

p1

```

```{r cit, echo=FALSE, message=FALSE, include=TRUE}

cit <- rem %>%
  filter(departed_date >= "2011-10-01",
       departed_date <= "2022-09-30") %>% 
  mutate(citizenship_country = toupper(citizenship_country)) %>% 
  group_by(citizenship_country) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

p1 <- rem %>% 
  filter(departed_date >= "2011-10-01",
       departed_date <= "2022-09-30") %>% 
  mutate(citizenship_country = case_when(
    citizenship_country %in% head(cit$citizenship_country, 15) ~ citizenship_country,
    TRUE ~ "all others"
  )) %>% 
  count(fy, citizenship_country) %>% 
  ggplot(aes(x=fy, y=n, fill=citizenship_country)) +
  geom_col() +
  labs(title = "Total ICE removals by country of citizenship (top 15)")

ggplotly(p1)
```

# ICE removals per AOR

```{r removals_per_aor_fy, echo=FALSE, message=FALSE, include=TRUE}

p1 <- rem %>% 
  filter(departed_date >= "2011-10-01",
         departed_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n, color=aor, group=aor)) +
  geom_line() +
  labs(title = "Total removals per FY by AOR")

ggplotly(p1)

```

Below is a chart of total ICE arrests per FY, highlighting the Seattle AOR ("SEA"):

```{r annual_sea}


p2 <- rem %>% 
  filter(departed_date >= "2011-10-01",
         departed_date <= "2022-09-30") %>% 
  group_by(fy, aor) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n, color=aor, group=aor)) +
  geom_line() +
  gghighlight(aor == "SEA") +
  labs(title = "Total removals per FY",
       subtitle = "Seattle AOR highlighted")

p2

```

```{r removals_pct_chg}

natl_pct_chg <- rem %>%
  filter(departed_date >= "2011-10-01",
         departed_date <= "2022-09-30") %>% 
  group_by(fy) %>%
  summarize(n = n()) %>% 
  mutate(pct_change = (n/lag(n) - 1))

p1 <- natl_pct_chg %>% 
  ggplot(aes(x = fy, y = pct_change)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(title="FY % change in total removals")

p1

aor_pct_chg <- rem %>%
  filter(departed_date >= "2011-10-01",
         departed_date <= "2022-09-30",
         aor != "HQ",
         !is.na(aor)) %>% 
  group_by(fy, aor) %>%
  summarize(n = n()) %>% 
  group_by(aor) %>% 
  arrange(fy, .by_group=TRUE) %>% 
  mutate(pct_change = (n/lag(n) - 1))

p2 <- aor_pct_chg %>% 
  ggplot(aes(x = fy, y = pct_change)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(breaks=seq(2012, 2022, 2)) +
  facet_wrap(~aor)  +
  labs(title="FY % change in total removals per AOR") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p2

# Could try superimposing line chart of total arrests as above

```

# Removals by processing disposition

```{r disp, echo=FALSE, message=FALSE, include=TRUE}

disps <- rem %>% 
  filter(departed_date >= "2012-10-01",
         departed_date <= "2022-09-30",
         ) %>% 
  count(processing_disposition_clean) %>% 
  arrange(desc(n))

top_disp <- disps %>% 
  filter(n > 50000)

rem <- rem %>% 
  mutate(disp_short = case_when(processing_disposition_clean %in% unlist(top_disp$processing_disposition_clean) ~ as.character(processing_disposition_clean), 
                                         TRUE ~ "ALL OTHERS"))


```

Fewer reinstatements of removal.

```{r rem_disp_fy}

p1 <- rem %>% 
  filter(departed_date >= "2012-10-01",
         departed_date <= "2022-09-30",
         ) %>% 
  group_by(fy, disp_short) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.factor(fy), y=n, fill=disp_short)) +
  geom_col() +
  labs(title = "Total removals per FY by processing disposition")

ggplotly(p1)

```

Decline of expedited removal.

```{r expedited_removal, echo=FALSE, message=FALSE, include=TRUE}

er <- c('er', 'ex')

rem <- rem %>% 
  mutate(processing_disp = tolower(processing_disp),
         ex_rem = case_when(processing_disp %in% er ~ 'expedited removal',
                            str_detect(processing_disp, 'expedited removal') ~ 'expedited removal',
                            is.na(processing_disp) ~ 'all others',
                            TRUE ~ 'all others'))

p1 <- rem %>% 
  group_by(year, ex_rem) %>%
  ggplot(aes(x = year, fill=ex_rem)) +
  geom_bar(stat='count')

p1

p1 <- rem %>% 
  group_by(year, aor, ex_rem) %>%
  ggplot(aes(x = year, fill=ex_rem)) +
  geom_bar(stat='count') +
  facet_wrap(~aor, scales="free_y")

p1

```

# Dates

Removals data includes four separate date fields. `departed_date` is most complete, with no missing values.

```{r date_compare, echo=FALSE, message=FALSE, include=TRUE}
rem %>% 
  dplyr::select(contains('date')) %>% 
skimr::skim()

hist(rem$departed_date, breaks='years', col='pink')
hist(rem$removal_date, breaks='years', col='lightblue')
hist(rem$arrest_date, breaks='years', col='lightyellow')
hist(rem$case_close_date, breaks='years', col='lightgreen')

```

We can calculate lag between different dates for example between departed date and removal date; unclear if this is particularly meaningful.

```{r dep_diff_rem}

rem$dep_diff_rem <- difftime(rem$departed_date, rem$removal_date, units='days')

median(rem$dep_diff_rem, na.rm=TRUE)
mean(rem$dep_diff_rem, na.rm=TRUE)
max(rem$dep_diff_rem, na.rm=TRUE)
min(rem$dep_diff_rem, na.rm=TRUE)

h1 <- hist(as.numeric(rem$dep_diff_rem))

p1 <- plot(as.numeric(rem$dep_diff_rem))

```

```{r arr_diff_rem}

rem$arr_diff_rem <- difftime(rem$arrest_date, rem$removal_date, units='days')

median(rem$arr_diff_rem, na.rm=TRUE)
mean(rem$arr_diff_rem, na.rm=TRUE)
max(rem$arr_diff_rem, na.rm=TRUE)
min(rem$arr_diff_rem, na.rm=TRUE)

h1 <- hist(as.numeric(rem$arr_diff_rem))

p1 <- plot(as.numeric(rem$arr_diff_rem))

```

```{r plot dates}

p1 <- plot(rem$id, rem$departed_date)

p2 <- plot(rem$id, rem$removal_date)

p3 <- plot(rem$removal_date, rem$departed_date)

p4 <- rem %>% 
  ggplot(aes(x = removal_date, y = departed_date)) +
  geom_point(aes(color = as.integer(departed_date)))

p4
```

